Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
references

Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].

@@

A robust understanding of forest impacts on global C cycling is essential. 
Total annual photosynthesis in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], more than seven times the average annual fossil fuel emissions during 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire [@van_der_werf_global_2017].
In recent years, total forest C uptake has exceeded releases, such that globally forests have been a C sink.
Considering only areas remaining in forest, this C sink has averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ for 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ [@friedlingstein_global_2019].

@@

###  Citations to add

@harris_global_2021

>>>>

Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
Large-scale synthesis is critical to benchmarking model performance with global data [@clark_field_2017; @luo_framework_2012], quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].

@@

A robust understanding of forest impacts on global C cycling is essential. 
Total annual photosynthesis in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], more than seven times the average annual fossil fuel emissions during 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire [@van_der_werf_global_2017].
In recent years, total forest C uptake has exceeded releases, such that globally forests have been a C sink [@harris_global_2021].
Considering only areas remaining in forest, this C sink has averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ for 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ [@friedlingstein_global_2019; @harris_global_2021].

@@
